---
title: "Research Methods in Developmental Linguistics -- Week 6"
subtitle: "Group-level effects: Slopes"
author: "Dr Stefano Coretta"
institute: "University of Edinburgh"
format:
  mono-light-revealjs:
    theme: [default, custom.scss]
    history: false
execute: 
  echo: false
---

```{r setup, include=FALSE}
library(tidyverse)
theme_set(theme_light())

library(brms)
library(tidybayes)

options(ggplot2.discrete.fill = RColorBrewer::brewer.pal(8, "Dark2"))
options(ggplot2.discrete.colour = RColorBrewer::brewer.pal(8, "Dark2"))
options(show.signif.stars = FALSE)

my_seed <- 8878
set.seed(my_seed)
options(scipen=999)
```


## Ota 2009: the study

- From *The KEY to the ROCK: Near-homophony in nonnative visual word recognition* ([10.1016/j.cognition.2008.12.007](https://doi.org/10.1016/j.cognition.2008.12.007))

- Visual semantic-relatedness paradigm.

  - LOCK/ROCK ~ KEY.

- Japanese data.

## Ota 2009: the data

```{r}
#| label: ota2009

ota2009 <- read_csv("data/ota2009/key-rock.csv") %>%
  filter(
    Procedure == "TrialProc", Contrast != "F"
  ) %>%
  mutate(
    Subject = as.factor(Subject), RT_log = log(Words.RT)
  )
ota2009
```

## Ota 2009: the data

```{r}
#| label: contr-cond

ota2009 %>%
  ggplot(aes(Contrast, RT_log)) +
  geom_violin(aes(fill = Condition), alpha = 0.2) +
  geom_jitter(
    aes(colour = Condition), alpha = 0.4,
    position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.9)
  ) +
  stat_summary(
    aes(group = Condition),
    position = position_dodge(width = 0.9),
    fun = "median", geom = "point", size = 3
  )
```

## Ota 2009: subject-level slopes

```{r}
#| label: subj-slo
#| fig_asp: 0.5

ota2009 %>%
  filter(Contrast == "LR") %>%
  ggplot(aes(Condition, RT_log)) +
  geom_violin(aes(fill = Condition), alpha = 0.5) +
  # geom_jitter(
  #   aes(colour = Condition), alpha = 0.4,
  #   position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.9)
  # ) +
  stat_summary(
    aes(group = Condition),
    position = position_dodge(width = 0.9),
    fun = "median", geom = "point"
  ) +
  facet_wrap(~ Subject, nrow = 2) +
  theme(axis.text.x = element_blank(), legend.position = "bottom")
```

## Ota 2009: intercepts and slopes

```{r}
#| label: ota-bm-3
#| echo: true

ota_bm_3 <- brm(
  RT_log ~
    Condition * Contrast +
    (Condition * Contrast | Subject),
  family = gaussian,
  data = ota2009,
  backend = "cmdstanr",
  seed = my_seed,
  file = "data/cache/ota_bm_3"
)
```

## Ota 2009: intercepts and slopes

```{r}
#| label: ota-bm-3-sdraws
#| echo: true

ota_bm_3_sdraws <- ota_bm_3 %>%
  spread_draws(`b_.*`, regex = TRUE) %>%
  mutate(
    LR_cond_dif = b_ConditionUnrelated + `b_ConditionUnrelated:ContrastLR`
  )
ota_bm_3_sdraws
```

## Ota 2009: difference between unrelated and control

```{r}
#| label: ota-bm-3-plot

ota_bm_3_sdraws %>%
  ggplot(aes(LR_cond_dif)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  stat_halfeye(slab_alpha = 0.5)
```

## Ota 2009: subject-level intercepts

```{r}
#| label: ota-bm-3-group
#| echo: true

ota_bm_3_group <- ota_bm_3 %>%
  gather_draws(r_Subject[Subject,term])
ota_bm_3_group
```

## Ota 2009: subject-level intercepts

```{r}
#| label: ota-bm-3-group-plot

ota_bm_3_group %>%
  filter(term == "ConditionUnrelated:ContrastLR") %>%
  ggplot(aes(.value)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  stat_halfeye() +
  facet_wrap(vars(Subject))
```

## Ota 2009: subject-level intercepts

```{r}
#| label: ota-bm-fixef
#| echo: false

ota_bm_1 <- readRDS("data/cache/ota_bm_1.rds")
ota_bm_2 <- readRDS("data/cache/ota_bm_2.rds")

fixef(ota_bm_1, probs = c(0.05, 0.95))
fixef(ota_bm_2, probs = c(0.05, 0.95))
fixef(ota_bm_3, probs = c(0.05, 0.95))
```

## Ota 2009: conditional effects (log-RT)

```{r}
#| label: ota-bm-3-cond

conditional_effects(ota_bm_3, "Contrast:Condition")
```

## Ota 2009: conditional effects (accuracy)

```{r}
#| label: ota-bm-acc

ota_bm_acc <- brm(
  Words.ACC ~
    Condition * Contrast +
    (Condition * Contrast | Subject),
  family = bernoulli,
  data = ota2009,
  backend = "cmdstanr",
  seed = my_seed,
  file = "data/cache/ota_bm_acc"
)
```

```{r}
#| label: ota-bm-acc-cond

conditional_effects(ota_bm_acc, "Contrast:Condition")
```

