---
title: "Research Methods in Developmental Linguistics -- Week 6"
author: "Dr Stefano Coretta"
institute: "University of Edinburgh"
format:
  mono-light-revealjs:
    theme: [default, custom.scss]
    history: false
execute: 
  echo: false
knitr:
  opts_chunk:
    dev: "ragg_png"
filters: ["tachyonsextra"]
---

## Song 2020

```{r}
#| label: setup
#| include: false

library(tidyverse)
theme_set(theme_light())

library(brms)
library(tidybayes)

options(ggplot2.discrete.fill = RColorBrewer::brewer.pal(8, "Dark2"))
options(ggplot2.discrete.colour = RColorBrewer::brewer.pal(8, "Dark2"))
options(show.signif.stars = FALSE)

my_seed <- 8878
set.seed(my_seed)
options(scipen=999)
```

::: box-note
-   Data from [Song 2020](https://doi.org/10.1017/S0272263120000170).

-   Shallow morphological representation in L2 speakers.

-   **Lexical decision task**: Is the word a real English word or not?

-   **Priming paradigm**:

    -   **Prime**: *prolong* (unrelated), *unkind* (constituent), *kindness* (non-constituent).

    -   **Target**: *unkindness* (*\[un-kind\]-ness*, not *un-\[kind-ness\]*).

-   English L1 and English L2/Cantonese L1 participants.
:::

## Song 2020: hypotheses

::: box-warning
Vague hypothesis:

> The Shallow Structure Hypothesis (SSH) allows the possibility for nonnative speakers to show strong priming effects from nonconstituent and constituent primes equally because the shallow processing route, which relies more heavily on semantic and surface-level information than on structural information.
:::

## Song 2020: the data

```{r}
#| label: shallow
#| echo: true

shallow <- read_csv("data/song2020/shallow.csv") |>
  filter(
    Critical_Filler == "Critical",
    Branching == "Left"
  ) |> 
  mutate(
    accuracy = ifelse(ACC == 0, "incorrect", "correct"),
    accuracy = factor(accuracy, levels = c("incorrect", "correct"))
  )
shallow
```

## Song 2020: accuracy

```{r}
#| label: shallow-acc

shallow |> 
  ggplot(aes(Relation_type, fill = accuracy)) +
  geom_bar(position = "fill") +
  facet_grid(cols = vars(Group))

```

## Bernoulli regression (accuracy)

```{r}
#| label: sha-bm-1
#| echo: true

my_seed <- 4235

sha_bm_1 <- brm(
  accuracy ~ 0 + Group:Relation_type,
  family = bernoulli,
  data = shallow,
  cores = 4,
  seed = my_seed,
  file = "data/cache/sha_bm_1"
)
```

## Bernoulli regression: expected predictions

```{r}
#| label: sha-bm-1-cond

conditional_effects(sha_bm_1, "Group:Relation_type")

```

## BUT...

::: box-error
-   Multiple observations from different participants.

-   Multiple observations from different target words.
:::

. . .

::: box-tip
-   We need to include **varying terms**.
:::

## By-subject accuracy

```{r}
#| label: shallow-id

shallow |> 
  filter(Group == "L2") |> 
  ggplot(aes(Relation_type, fill = accuracy)) +
  geom_bar(position = "fill") +
  facet_wrap(~ID)
```

## By-target accuracy

```{r}
#| label: shallow-target

shallow |> 
  filter(Group == "L2") |> 
  ggplot(aes(Relation_type, fill = accuracy)) +
  geom_bar(position = "fill") +
  facet_wrap(~Target)
```

## By-subject and by-target varying terms

```{r}
#| label: sha-bm-2
#| echo: true

sha_bm_2 <- brm(
  ACC ~ 0 + Group:Relation_type +
    (0 + Relation_type | ID) +
    (0 + Group:Relation_type | Target),
  family = bernoulli,
  data = shallow,
  cores = 4,
  seed = my_seed,
  file = "data/cache/sha_bm_2"
)
```

## By-subject and by-target model: expected predictions plot

```{r}
#| label: sha-bm-2-conds
conditional_effects(sha_bm_2, "Group:Relation_type")
```

## Difference between non/constituent in L2

```{r}
#| label: sha-bm-2-draws
#| echo: true

sha_bm_2_draws <- as_draws_df(sha_bm_2, variable = "b_", regex = TRUE)

sha_bm_2_draws <- sha_bm_2_draws |> 
  mutate(
    diff_l2 = `b_GroupL2:Relation_typeNonConstituent` - `b_GroupL2:Relation_typeConstituent`
  )
```

## Difference between non/constituent in L2: plot

```{r}
#| label: sha-bm-diff-plot

sha_bm_2_draws |> 
  ggplot(aes(diff_l2)) +
  geom_vline(xintercept = 0) +
  stat_halfeye()
```

## Difference between non/constituent in L2: CrIs

```{r}
#| label: sha-bm-diff-diff-cri
#| echo: true

sha_bm_2_draws |> 
  select(diff_l2) |> 
  summarise_draws(~quantile(.x, probs = c(0.05, 0.95, 0.5, 0.85))) |> 
  mutate(
    across(where(is.numeric), ~round(.x, digits = 2))
  )
```

## Results overview

::: box-note
> The Shallow Structure Hypothesis (SSH) allows the possibility for nonnative speakers to show strong priming effects from nonconstituent and constituent primes equally because the shallow processing route, which relies more heavily on semantic and surface-level information than on structural information.

- The 90% CrI of the difference between nonconstituent and constituent in L2 is [-1.3, 0.26].

- At 70% probability, the log-odds of the priming effect on accuracy between nonconstituent and constituent in L2 participants differs between -0.01 and -0.47.

:::

## Summary

::: box-tip
-   You should include varying terms if data is "hierarchical", for example repeated measures from participants or items.

- Not including varying terms (wrongly) inflates posterior certainty.

- Frequentist regression models with lme4 often don't converge and researchers simplify the hierarchical structure of the model.
:::
